Friends Get More Money, Attention and Handshakes: Chinese Foreign Aid, Xinhua and Diplomatic Visits

### 2022-09-07 04_paper/Draft4_812.Rmd
trying

However, the discovery that China is little different from traditional Western donors makes the limited state-sanctioned transparency of its overseas development financing all the more puzzling. Many of the goals often attributed to Western donors (generating goodwill, strengthening diplomatic ties, etc.) are difficult to accomplish in secret. And, while Beijing has taken steps to improve the visible branding of its aid projects on the ground, there is limited evidence that this is sufficient for attributing credit to Beijing [see @dietrichEtAl2018]. This therefore raises a natural question. *Do China’s overseas aid and debt relate to public diplomacy and international recognition? If so, how?*

@@

While public diplomacy and outward-facing state-sponsored media are not solely used to support China's overseas development financing, we believe they have clear overlapping purposes. In particular, diplomacy and coverage are critical *visible* activities that can complement the often opaque character of Beijing's development finance. Thus, to the extent that China's development finance is targeted at supporting image- and alliance-building activities, we hypothesize that recipients of China's overseas development assistance will also be disproportionate targets of media coverage in *Xinhua* (to support China's image and to build publicity around its development activities) and be more frequent hosts of diplomatic missions from Beijing (to support closer bilateral diplomacy and cooperation).

As we test these hypotheses, in our research design we take care to ensure that we delineate China's development financing according to types of assistance China offers. Until recently, many have mistakenly conflated Beijing's ODA-like giving with its other official financing (OOF). However, efforts to disentangle China's activities with respect to grants and low-interest loans (financing that would meet the OECD definition of official development assistance or ODA) from its activities with respect to higher interest loans and export credits (OOF) have unmasked a considerable divergence in goals. The conventional view is that while *aid* (ODA) is closely tied to Beijing's foreign policy goals, *debt* (OOF) is related to Beijing's commercial interests.^[See @dreherEtAl2022.] Thus, as we test our hypotheses, we also distinguish between Beijing-as-donor from Beijing-as-lender.

@@

>>>>

However, the discovery that China is little different from traditional Western donors makes the limited state-sanctioned transparency of its overseas development financing all the more puzzling. Many of the goals often attributed to Western donors (generating goodwill, strengthening diplomatic ties, etc.) are difficult to accomplish in secret. And, while Beijing has taken steps to improve the visible branding of its aid projects on the ground, there is limited evidence that this is sufficient for attributing credit to Beijing [see @dietrichEtAl2018]. This therefore raises a natural question. *Does China’s overseas development financing relate to public diplomacy and international recognition? If so, how?*

@@

While public diplomacy and outward-facing state-sponsored media are not solely used to support China's overseas development financing, we believe they have clear overlapping purposes. In particular, diplomacy and coverage are critical *visible* activities that can compensate for the often opaque character of Beijing's development finance. Thus, to the extent that China's development finance is targeted at image-building and diplomatic activities, we hypothesize that recipients of China's overseas development assistance will also be disproportionate targets of media coverage in the English version of *Xinhua* (to support China's image and to build publicity around its development activities) and be more frequent hosts of diplomatic missions from Beijing (to support closer bilateral diplomacy and cooperation).

As we test these hypotheses, in our research design we take care to ensure that we delineate China's development financing according to the types of assistance China offers. Until recently, many have mistakenly conflated Beijing's ODA-like giving with its other official financing (OOF). However, efforts to disentangle China's activities with respect to grants and low-interest loans (financing that would meet the OECD definition of official development assistance or ODA) from its activities with respect to higher interest loans and export credits (OOF) have unmasked a considerable divergence in goals. The conventional view is that while *aid* (ODA) is closely tied to Beijing's foreign policy goals, *debt* (OOF) is related to Beijing's commercial interests.^[See @dreherEtAl2022.] Thus, as we test our hypotheses, we also distinguish between Beijing-as-donor from Beijing-as-lender.

@@

Figure 1 shows total ODA (aid) and OOF (debt) disbursed by Beijing per year from 2002 to 2017. Values are shown in millions of 2017 USD. Blue shows values for aid and red shows values for debt. The difference in China's application of debt relative to aid is stark, and the gulf only expands over time. The observed explosion in debt relative to aid aligns with a shift in Beijing's strategy going into the twenty-first century, as others have documented. Burdened with vast foreign currency reserves due do years-long trade surpluses, confronted with excess domestic production of industrial goods, and in need of access to resources to support its continued economic growth, Beijing tasked its so-called policy banks to help Chinese firms gain access to foreign markets [@dreherEtAl2022]. This focus spurred a radical change in China's development finance portfolio. As @dreherEtAl2022 put it, "Beijing began to behave less like a benefactor and more like a banker" (5). Figure 1 shows this transformation from donor to lender in the starkest of terms.
